#**Dimensionality and Reduction Techniques**#

For this exercise I use the newly created dataset *human*, which is based on the two datasets *hd* and *gii* from the *Human Development Index*. The original datasets consisted of 195 observations each, the hd-dataset originally included 7 observations, while the gii-dataset included 10 variables.

Four variables of the dataset were reduced to the number of two by combining the separate variables  Edu2.F and  Edu2.M to Edu2.FM and the separate variables to Lab.FM.

In order to create the dataset *human* the two sets have been joined by the variable *country*. The final seven observations have been removed, as they referred to the regions World, Sub-Saharan Africa, South Asia, Latin America and the Caribbean, Europe and Central Asia, East Asia and the Pacific, and Arab States, and as such not to specific countries. 

As can be seen below the dataset *human* now consists of the following 8 variables:

- **Life_Exp** refers to average life-expectency in the number of years from the date of birth
- **Exp_y_Edu** refers to the expected number of educational years
- **GNIpCap** refers to the Gross national income (GNI) per
capita
- **MatMorRatio** refers to the Maternal mortality ratio	  
- **Adol_Birth_Rate** refers to the Adolescent birth rate	
- **ParlRepPercent** refers to the women's share of seats in parliament
- **Edu2.FM** refers to the population with at least some secondary education for both males and females of the age of at least 25)
- **Lab.FM** refers to the Labour force
participation rate (for both males and females of the age of 15 or older)

And the dataset itself now consists of 155 observations.

Accessing libraries and loading the dataset "human":
```{r libraries and loading human-dataset}


#activating libraries:
library(tidyr)
library(corrplot)
library(ggplot2)
library(dplyr)

#loading human:
human <- read.csv("C:\\Users\\Ani\\Documents\\GitHub\\IODS-project\\data\\human.csv", header=TRUE, row.names = 1)

```

Here we take a quick look at the structure and dimensions of the dataset, which are sorted by the name of the country:
```{r human structure & data}

str(human)
dim(human)

```

Let's take a brief look at the individual variables of the dataset "human":

```{r human variable summaries}

summary(human$Life_Exp)
summary(human$Exp_y_Edu)
summary(human$GNIpCap)
summary(human$MatMorRatio)
summary(human$Adol_Birth_Rate)
summary(human$ParlRepPercent)
summary(human$Edu2.FM)
summary(human$Lab.FM)
```

As I had some trouble that I could not solve in my attempts I have **excluded the variable "GNIpCap" from being visualized in the following graph**. 
I am aware that it was not part of the task, but even after asking for help in the course-session, I could not get this to work without RStudio giving me a warning about the "cardinality threshold". 

```{r graphical visualization}

library(GGally)
human <- select(human, -GNIpCap)
ggpairs(human, cardinality_threshold = NULL)
cor(human)%>%corrplot

```

