

# **Week 2: Let's start with the data**

*Describe the work you have done this week and summarize your learning.*

## 1. Data at a glance##
After reading the data into R, I took a glimpse at the provided data. Here are my first observations:

1. The data-set consists of 183 observations relating to 60 variables.
2. Except for the variables "Age", "Attitude", "Points", and "gender" the data appear to be provided by a scale from 1 to 5. (In the print out, however, 167 rows were omitted, thus a higher scale may be possible)
3. The variable "gender", as a qualitative characteristic, are given as binary text-data in "M" (=male) and "F" (=female).
4. The variable "Age", is provided as number of years, and derived from the provided year of birth by the respondent. 

## 2. Data-Wrangling Attempts ##

**1. I printed and summarized the variable "gender":**

  - the result showed that 122 respondents were female, and 61 were male.
  
**2. I printed and summarized the variable "Age".**

  - The age-range is between 17 and 55 years.
  - The average age is 25,58 years.
  - The median-age is 22 years.
  
**3. I printed and summarized the variable "Points"**
 
  - At this point the variable ranges between 0 and 33.
  - The average points are 20,61.
  - The median-value is 22,00 points.
  
**4. I scaled down the variable "Attitude" by 10, afterwards I printed and summarized it**
 
  - The variable - according to the additional information homepage - intends to reflect the respondent's "global attitude towards statistics"
  - The range of the variable-values is now between 1,4 and 5,00. 
  - The average-value is 3,121.
  - The Median-value is 3,2.
  
**5. I attempted to establish the combination variables "deep_questions", "surface_questions", and "strategic_questions"**

  - I combined the variables for each as it was done in the datacamp-exercises ("deep_questions" and "surface_questions" thus resulted out of a combination of 12 variables each, while "strategic_questions" became a combination of 8 variables).
  - Next I created the columns out of the individual new variables and named them "deep_columns", "surface_columns", and "startegic_columns".
  - I attemped to apply the rowMeans function on each. While in the print outs it showed me the columns results, I could not get the summarise function to work. I am not sure if that is supposed to be that way.
  
**6. I created the new data-set "learning2014"**

  - At this point it consisted of 183 observations and 7 variables.
  - The variable names are "gender", "Age", "Attitude", "deep", "stra", "surf", and "Points".
  - After I tried to exclude respondents with 0 points within the "Points"-variable my dataset consisted of only 166 observations for the 7 given variables. 

Up to this point, it appears to me as though it worked according to the instructions...


##Trying to establish the .csv-file##

It appears, that I have managed to create the file, but I am not sure if I manage any further... Here, I guess, you will be able to see why I am not happy with it...

```{r }

learning2014 <- read.csv("C:\\Users\\Ani\\Documents\\GitHub\\IODS-project\\data\\learning2014.csv", header=TRUE)

```

Looking at the structure of my dataset:

```{r dim}

dim(learning2014)

```

*I am not sure why the dim-function tells me here that I have 8 variables and not seven as in the r-file...*

Trying to explore the structure of my data:
```{r str}

str(learning2014)

```
With the exception of this extra-variable which I cannot explain - I'm sorry - it appears that otherwise the 7 variables "gender", "Age", "Attitude", "deep", "surf", stra", and "Points" are there. I furthermore appear to be having the 166 observations. Let's see how it may, or may not work out with the rest of the tasks...

- Describe your work and results clearly.
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  
