#**Week 3: Logistic Regression**#

##Data at a glance##

####The original Datasets:####

The data-set used here is based on the datasets provided by the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Student+Performance).

Citation of the dataset-reference:
*P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7.*

The original datasets "student_mat.csv" (including 395 observations and 33 variables, with regard to the subject of mathematics) and "student_por.csv" (consisting of 649 observations and 33 variables, relating to the subject of Portuguese) were established from data collected at two Portuguese secondary schools.

The data attributes of these original datasets include student grades, demographic, social and school related features) and it was collected by using school reports and questionnaires.

The three final variables "G1", "G2", and "G3" relate to the specific grades the respondent has achieved either in mathematics or in Portuguese, while the other variables operationalize more general conditions of the respective respondent, relating to his or her familial background, school- and leisure-time habits, and general information (i.e., for instance, age and gender).


####The altered dataset used in this project:####

In order to look at my data-wrangling coding for this week's exercise, feel free to [go this way](https://github.com/JoaWe/IODS-project/blob/master/data/create_alc.R).

The two datasets were joined by means of using the variables "school", "sex", "age", "address", "famsize", "Pstatus", "Medu", "Fedu", "Mjob", "Fjob", "reason", "nursery","internet" as identifiers. 
Only the respondents present in both datasets were kept for the newly established dataset.

Two new variables, "alc_use" and "high_use", have been established in the new dataset, by the following means:

1.**alc_use**: The average of the answers related to weekday and weekend alcohol consumption have been taken.

2.**high_use**: Then the newly created variable "alc_use" has been utilized to create a second new variable "high_use", which is TRUE for students for whom "alc_use"" is greater than 2.

Accordingly, this new dataset now contains **382 observations relating to 35 variables**.


- Reading the joined student alcohol-consumption data into R from my data-folder:

```{r }

alc <- read.csv("C:\\Users\\Ani\\Documents\\GitHub\\IODS-project\\data\\alc.csv", header=TRUE)

```

- Summarizing my data-set:

```{r summary}

summary(alc)

```


- the dimensions of the dataset:
```{r dim}

dim(alc)

```

- Listing the column-names, to show the variables:

```{r colnames}

colnames(alc)

```
