#**Clustering and Classification**#

####Loading and exploring the data *Boston*:####


```{r loading and exploring "Boston"}


#activating libraries
library(tidyr)
library(corrplot)
library(ggplot2)
library(dplyr)

#access MASS-Package
library(MASS)

#load the data
data("Boston")

#exploring the data
dim(Boston)
str(Boston)
summary(Boston)

```

The dataset consists of 14 variables including 506 observations relating to the city of Boston.

####Graphical Overview of the dataset:####


####Scaling the dataset "Boston"####

```{r standardizing "Boston"}

#center and standardize "Boston"
boston_scaled <- scale(Boston)

#summarizing boston scaled
summary(boston_scaled)

#changing the object to a dataframe
boston_scaled <- as.data.frame(boston_scaled)

```

####Creating Variable "crime"####

In the following steps I intend to create a new categorical variable, named "crime", and use it to replace the original variable "crim".
The quantiles are used as break points for this categorical variable. Finally the new dataset will be divided to train- and test-sets, so that 80% of the data belongs to the train set.

```{r creating variable "crime"}

#summary of scaled variable "crim"
summary(boston_scaled$crim)

#creating and printing the quantile vector of "crim"
bins <- quantile(boston_scaled$crim)

#printing quantile vector
bins

#creating categorical variable "crime"
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))

#creating table "crime"
table(crime)

#removing original variable "crim" from the dataset
boston_scaled <- dplyr::select(boston_scaled, -crim)

#adding new variable to scaled dataset
boston_scaled <- data.frame(boston_scaled, crime)

#number of rows in the "Boston" dataset
n <- nrow(boston_scaled)

#choosing 80% of the rows
ind <- sample(n, size = n * 0.8)

#creating the train-set
train <- boston_scaled[ind,]

#creating test-set
test <- boston_scaled[-ind,]

#saving the correct classes from test-data
correct_classes <- test$crime

#removing the variable "crime" from the test-data
test <- dplyr::select(test, -crime)

```

