#**Week 3: Logistic Regression**#

##Data at a glance##

####The original Datasets:####

The data-set used here is based on the datasets provided by the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Student+Performance).

Citation of the dataset-reference:
*P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7.*

The original datasets "student_mat.csv" (including 395 observations and 33 variables, with regard to the subject of mathematics) and "student_por.csv" (consisting of 649 observations and 33 variables, relating to the subject of Portuguese) were established from data collected at two Portuguese secondary schools.

The data attributes of these original datasets include student grades, demographic, social and school related features) and it was collected by using school reports and questionnaires.

The age range of the respondents lies between 15 and 22 years.

The three final variables "G1", "G2", and "G3" relate to the specific grades the respondent has achieved either in mathematics or in Portuguese, while the other variables operationalize more general conditions of the respective respondent, relating to his or her familial background, school- and leisure-time habits, and general information (i.e., for instance, age and gender).


####The altered dataset used in this project:####

In order to look at my data-wrangling coding for this week's exercise, feel free to [go this way](https://github.com/JoaWe/IODS-project/blob/master/data/create_alc.R).

The two datasets were joined by means of using the variables "school", "sex", "age", "address", "famsize", "Pstatus", "Medu", "Fedu", "Mjob", "Fjob", "reason", "nursery","internet" as identifiers. 
Only the respondents present in both datasets were kept for the newly established dataset.

Two new variables, "alc_use" and "high_use", have been established in the new dataset, by the following means:

1.**alc_use**: The average of the answers related to weekday and weekend alcohol consumption have been taken.

2.**high_use**: Then the newly created variable "alc_use" has been utilized to create a second new variable "high_use", which is TRUE for students for whom "alc_use"" is greater than 2.

Accordingly, this new dataset now contains **382 observations relating to 35 variables**.


- Reading the joined student alcohol-consumption data into R from my data-folder:

```{r }

alc <- read.csv("C:\\Users\\Ani\\Documents\\GitHub\\IODS-project\\data\\alc.csv", header=TRUE)

```

- structure of my data-set:

```{r str_alc}

str(alc)

```


- the dimensions of the dataset:
```{r dim_alc}

dim(alc)

```


##Hypotheses relating to students' alcohol-consumption##

Accessing libraries:
``` {r accessing libraries}

library(tidyr) 
library(dplyr) 
library(ggplot2)

```


With regard to higher or lower alcohol-consumption among students certain hypotheses can be presumed relating to the following **four variables** taken from the newly established dataset "alc.csv". These hypotheses are to be tested and (dis-)proven here.
The dependent variables of interest are...

...**"alc_use"**, which displays the average-answers related to weekday and weekend alcohol consumption (from 1 = very low consumption to 5 = very high consumption) . For a brief glimpse at the respective variable, see the barplot here:

```{r alc$alc_use & barplot}

g1 <- ggplot(data = alc, aes(x = alc_use))

g1 + geom_bar()

```



...and **"high_use"**, which is a logical column based on the variable *"alc_use"*. High usage of alcohol-consumption is therein defined asbeing TRUE, if the observation in the alc_use-variable is greater than 2. Take a sneak-peak at the summary of this variable: 

```{r alc$high_use & barplot}

summary(alc$high_use)

g2 <- ggplot(alc, aes(high_use))

g2 + facet_wrap("sex") + geom_bar()

```

```


1. It can be assumed that a respective student's ambitions can impact on his or her alcohol consumption habits. Accordingly, it is hypothesized that there is a negative correlation between the students' alcohol-consumption and their interest in pursuing a higher education degree.

       - **independent variable "higher":** reports on whether - or not, the respondent aims for a higher education degree, in a binary fashion of "yes" or "no". The summary of this variable follows here:

```{r summary_alc$higher}

summary(alc$higher)

```
     
An interesting observation at this point is that only 18 out of the 382 respondents of this dataset are *not* interested in achieving a higher educational degree.


2. Presuming that the time a student has to spare impacts on his or her alcohol-consumption level, two variables are considered to be of interest:

      a) The amount of time a student invests on studying ought to impact on the time a student has to, for instance, consume alcohol; meaning that the more time a student brings up to study the lower should be his or her alcohol consumption
      
      -**independent variable "studytime":** the observations here relate to the amount of studytime that a respondent estimates to be using per week, and they are presented numerically (1 = less than 2 hours; 2 = 2 to 5 hours; 3 = 5 - 10 hours; 4 = more than 10 hours ). *A summary of this variable will not be presented here, as due to the nature of the variable's data, a summary would be missleading.*


      b) A student's time can also be affected by whether, or not, that student uses his or her time for extra-curricular activities; meaning that a student that spends his or her time on extra-curricular activities, ought to display a lower alcohol-consumption.
      
      -**independent variable "activities":** The variable refers to the question of whether, or not, a student partakes in extracurricular activities, and it is thus represented in a binary fashion of "yes" or "no". 
            
      *The continuation-question that would require the respondents to estimate, how much time, and to report, for example, on the diversity of the respective activities is regrettably not represented in the dataset.*
      
```{r summary_alc$activities}

summary(alc$activities)

```

3. The final assumption, to be tested here, concerns whether the familial background, can be considered to play any role in the students' alcohol-consumption habits. There are several variables included in the dataset that could be drawn on for this hypothesis (e.g.: "famsize", "Medu", or "Fedu"):

    -**independent variable "Pstatus":** The variable depicts the the parents cohabitation-status in a binary fashion, in which "T" represents the parents living together, while "A" represents the parents living apart:

```{r summary_alc$Pstatus}

summary(alc$Pstatus)

```
      
